Assessing primitives performance\\ on multi-stage execution

### 2018-05-09 paper.tex
comparison, appendix

\usepackage{xcolor}
\newcommand{\todo}[1]{\color{orange}\fbox{\bfseries\sffamily\scriptsize TODO:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sd}[1]{\color{red}\fbox{\bfseries\sffamily\scriptsize Stef:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sk}[1]{\color{blue}\fbox{\bfseries\sffamily\scriptsize Sophie:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\cba}[1]{\color{purple}\fbox{\bfseries\sffamily\scriptsize Clement:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand*\rot{\rotatebox{90}}
\input{macros}

@@

 \todo{Finish section}

 In this section we compare the five different implementations according to the following criteria:
 \begin{itemize} 
 	\item \emph{Recompilation needed:} Diff no, plugin, full
	\item \emph{Implementation language familiar to developers:} Merged the 2 latter, explain ST dev ST, VM dev ST-Slang-C-ASM, maybe result is + ou - ? or + or 0
	\item \emph{Readability:} explain why number of lines of code. 
	\item \emph{Number of implementations to maintain:} explain fall-back code which is in all languages, Slang/C, JIT => maybe we ignore the fall-back code ?
	\item \emph{Stack switch from generated native code:} 
	\item \emph{Optimization potential:} discuss inlining C comp cross file and InterpreterProxy indirection
	\item \emph{Control over the generated code:} diff llvm gcc while JIT always the same
\end{itemize}

\begin{table}
\centering
\begin{tabular}{lccccccc}
  \hline
    & \rot{recompilation needed} & \rot{familiar to client dev} & \rot{familiar to vm dev} & \rot{number of LoC} & \rot{Implementations to maintain} & \rot{Performance} & \rot{Control over generated code} \\
  \hline
  Baseline    &             & \checkmark  &            & 10  & 3+ &  +  &  \\
   \hline
  SmartSyntax &             &  \checkmark &            & 20  & 3+ & ++  & -\\
   \hline
  SlangPlugin &             &  \checkmark &            & 35  & 2  & ++  & - \\
   \hline
  PureSlang   & \checkmark  &             & \checkmark & 45  & 1  & ++  & -\\
   \hline
  Slang+JIT   & \checkmark  &             & \checkmark & 85  & 1  & ++++ & + \\
  \hline
\end{tabular}
\caption{Trade-offs between implementations: summarized}
\label{tab:cmp_implem}
\end{table}

@@

@@

\begin{figure*}[htp!]
		\centering
		\includegraphics[width=1\linewidth]{figures/asm_cmp.pdf}
		\caption{GCC and LLVM: different optimizations on the comparison loop}
		\label{fig:Asm}
\end{figure*}

@@

%% Appendix
\vfill\null %switch column
\appendix
\section{Smart Syntax code}
\label{SmartSyntaxAppendix}

>>>>

\usepackage{xcolor}
\newcommand{\todo}[1]{\color{orange}\fbox{\bfseries\sffamily\scriptsize TODO:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sd}[1]{\color{red}\fbox{\bfseries\sffamily\scriptsize Stef:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sk}[1]{\color{blue}\fbox{\bfseries\sffamily\scriptsize Sophie:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\cba}[1]{\color{purple}\fbox{\bfseries\sffamily\scriptsize Clement:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
%\newcommand*{rotatebox{75}}
\input{macros}

@@

\begin{table*} [bth]
\centering
\begin{tabular}{lccccccc}
  \hline
    & Recompilation & Familiar to & Lignes & Number of & Stack switch & Optimization & Code generation \\
    & required & developers & of code & implementations & overhead & potential & control \\
  \hline
  Baseline    & No & All  & 10  & 0 & No & - & - \\
   \hline
  SmartSyntax & Plugin & VM & 20  & 3+ & Yes  & Limited & Low \\
   \hline
  SlangPlugin & Plugin & VM & 35  & 1  & Yes & Limited & Low \\
   \hline
  PureSlang   & Yes  & VM & 45  & 1  & Yes & High & Low \\
   \hline
  Slang+JIT   & Yes  & VM-JIT & 85  & 2  & No & High & Full \\
  \hline
\end{tabular}
\caption{Comparison between implementations\vspace{-0.5cm}}
\label{tab:cmp_implem}
\end{table*}

 In this section we compare the five different implementations according to the following criteria:
 \begin{itemize} 
 	\item \emph{Recompilation required:} Changing the implementation requires only Smalltalk bytecode recompilation (No), requires to recompile a plugin to a dynamic library if compiled as external or the recompilation of the full VM if compiled as internal (-) or requires to recompile the full VM (Yes).
	\item \emph{Implementation language familiar to developers:} Smalltalk application developers are considered to be familiar with Smalltalk only, while Cog VM developers are considered being familiar with Smalltalk, Slang, C and Assembly code. Only Cog's JIT implementors are familiar with its RTL, which is a subset of the VM developers.  We use \emph{All} if all communities are familiar with the language, \emph{VM} if all VM developers are and \emph{VM-JIT} if only a subset of VM developers are. 
	\item \emph{Readability:} To evaluate the readability of the approach, we use one of the most used metric, the number of lines of code.
	\item \emph{Number of implementations to maintain:} In each case, the client language has to implement a fall-back code in Smalltalk to execute if the primitive fails. This code is maintain by the client language implementors, and not by the VM implementors. Hence we do not count the fall-back code in the number of implementation to maintain, nor the pure Smalltalk implementation. However, as a VM implementor, we need to maintain what generates the C code of the VM, hence, in the case of Smart Syntax, we need to maintain a different method in each client language. We also count all the Slang implementations as well as the ones on Cog's RTL.
	\item \emph{Stack switch overhead:} Yes if activating the primitive from the native code generated by the JIT requires to switch between the C and the client stack, inducing execution time overhead.
	\item \emph{Optimization potential:} \todo{}
	discuss inlining C comp cross file and InterpreterProxy indirection
	Smalltalk we control indirectly.
	\item \emph{Control over the generated code:} \todo{}diff llvm gcc while JIT always the same
\end{itemize}

@@

\begin{figure*}[bth]
		\centering
		\includegraphics[width=1\linewidth]{figures/asm_cmp.pdf}
		\caption{GCC and LLVM: different optimizations on the comparison loop}
		\label{fig:Asm}
\end{figure*}

@@

@@

\vspace{10cm} %Adapt to make vfill work

%% Appendix
\null \vfill\null %switch column

\appendix
\section{Smart Syntax code}
\label{SmartSyntaxAppendix}
